---
title: "Product Sales Analysis"
author: "Akiel Walsh"
date: "2023-03-26"
output: pdf_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(SnowballC)
library(stringr)
library(ggplot2)
library(wordcloud)
library(RColorBrewer)
library(dplyr)
library(httpuv)
library(openssl)
library(RJSONIO)
library(tidyr)
library(wordcloud2)
library(tidytext)
library(lubridate)

```

```{r}
sales.df <- read.csv(file = "/home/walsh/Documents/Project/sales_data_2023.csv", stringsAsFactors = FALSE)
```

#Data Exploration of the Sales Dataset
```{r}
head(sales.df)
glimpse(sales.df)
str(sales.df)
```

```{r}
sales.df$Order.Date <- as.Date(sales.df$Order.Date, format = "%m/%d/%Y")
sales.df$Ship.Date <- as.Date(sales.df$Ship.Date, format = "%m/%d/%Y")
sales.df$Units.Sold <- as.numeric(sales.df$Units.Sold)
sales.df$Unit.Price <- as.numeric(sales.df$Unit.Price)
sales.df$Unit.Cost <- as.numeric(sales.df$Unit.Cost)
sales.df$Total.Revenue <- as.numeric(sales.df$Total.Revenue)
sales.df$Total.Cost <- as.numeric(sales.df$Total.Cost)
sales.df$Total.Profit <- as.numeric(sales.df$Total.Profit)
str(sales.df)
```

```{r}
#Viewing the different values for region, in cleaning we might just opt to replace these values or fully remove them
unique(sales.df$Region)
unique(clean_df$X)
```

```{r}
#Total # of Rows with no value or empty strings as well as columns with arbitrary values
total_empyt_rows <- sum(sales.df$Region == "" & sales.df$Country == "" & sales.df$X == "")
incorrect_vals <- sum(sales.df$Region == "?" | sales.df$Region == "???" | sales.df$Region == "k" | sales.df$Region == "\"\"" | sales.df$Region == "??" )

print(paste("Total Empty Rows is = " , total_empyt_rows , "Total Rows with incorrect Values = " , incorrect_vals))
```

```{r}
colm_names <- names(sales.df)
invalidVals <- c("", "?", "???", "k", "\"\"", "??", NA, "None")
invalidVal <- c( "")

for (colm in colm_names) {
  print(paste("Invalid or null values in column", colm, ":", sum(sales.df[[colm]] %in% invalidVal)))
}

## We can see the amount of empty or null values within the data set, we will remove these later.  
```


# Histograms for the different attributes of the dataset
```{r}
for (colm in colm_names){
  if ((class(sales.df[[colm]])) == "numeric"){
    hist(sales.df[[colm]], main = paste("Histogram of ", colm, sep = ""))
  }
}
```

# Boxplot to Identify Outliers
```{r}
for (colm in colm_names){
  if ((class(sales.df[[colm]])) == "numeric"){
    boxplot(sales.df[[colm]], main = paste("Boxplot of ", colm, sep = ""))
  }
}
```
#From the Results we can see that Total.Profit, Total Costs as well as Total Revenues contain a large number of outliers. We will adjust to fall within the median after we have removed the necessary rows and tables



#CLEANING DATASET
##FIXING MISSING NUMERIC VALUES BY REPLACING THEM WITH THE MEDIAN
```{r}
imputedVals.df <- sales.df

for (colm in colm_names) {
  if (class(sales.df[[colm]]) == "numeric") {
    col_median <- median(sales.df[[colm]], na.rm = TRUE)
    imputedVals.df[[colm]][is.na(sales.df[[colm]])] <- col_median
  }
}
```

## Removing invalid values and Nulls
```{r}
clean_df <- imputedVals.df
for (col in colm_names) {
  clean_df <- clean_df[!(clean_df[[col]] %in% invalidVals | is.na(clean_df[[col]])), ]
}
```
## Verifying  if data was cleaned
```{r}
for (colm in colm_names) {
  print(paste("Invalid or null values in column", colm, ":", sum(clean_df[[colm]] %in% invalidVals)))
}
```

```{r}

#sdf <- clean_df

for (colm in colm_names){
  if (class(clean_df[[colm]]) == "numeric") {
  col_median <- median(sdf[[colm]], na.rm = TRUE)
  sdf[[colm]][sdf[[colm]] %in% boxplot(sdf[[colm]])$out] <- col_median
  }
}

boxplot(sdf$Total.Profit)
boxplot(sales.df$Total.Profit)
boxplot(sdf$Units.Sold)

  
```



```{r}
sdf <- clean_df

# Compute outliers for each column
outliers <- sapply(clean_df, function(x) boxplot.stats(x)$out)

# Replace outliers with column medians
for (colm in colm_names){
  col_median <- median(clean_df[[colm]], na.rm = TRUE)
  sdf[[colm]][sdf[[colm]] %in% outliers[[colm]]] <- col_median
}
```

#Data Analysis 
##a. What is the average amount spent on Personal Care vs. Cosmetics in each region? [4]
```{r}
#a. What is the average amount spent on Personal Care vs. Cosmetics in each region? [4]
comparison.df <- sdf %>% select(X, Region, Total.Revenue)%>% 
  group_by(Region, X) %>% 
  filter(X == "Personal Care" | X == "Cosmetics") %>%
  summarise(mean_revenue = mean(Total.Revenue), .groups = "drop")  
```

```{r}
#Change Structure of code
ggplot(comparison.df, aes(x = Region, y = mean_revenue, fill = X)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Mean Revenue by Region and Category", x = "Region", y = "Mean Revenue") +
  scale_fill_manual(values = c("Personal Care" = "#5eb56b", "Cosmetics" = "#855eb5")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


##b. In which country do people use the most cosmetics?
```{r}
partb.df <- sdf %>% select(Country,Units.Sold, X) %>% filter(X == "Cosmetics")

partb.df <- partb.df %>% group_by(Country) %>% summarise(Max.Units = max(Units.Sold))
resultb <- partb.df %>% select(everything()) %>% filter(Max.Units == max(Max.Units))

print(paste("The country that uses the Most Cosmetic is ", resultb$Country , "with the total number of units sold being : ", resultb$Max.Units))

```

##c In 2016, which product had the highest and lowest quantity sold across each region?

```{r}

partc.df <- sdf %>% 
  select(Region, Units.Sold, X, Order.Date) %>% 
  filter(as.Date(Order.Date) > as.Date('2016-01-01') & as.Date(Order.Date) <= as.Date('2016-12-31'))

partcmax.df <- partc.df %>% group_by(Region) %>% 
  #summarise(max.sold = max(Units.Sold), min.sold = min(Units.Sold))
  summarise(max_product = X[which.max(Units.Sold)], 
            max_sold = max(Units.Sold),
            min_product = X[which.min(Units.Sold)],
            min_sold = min(Units.Sold))
View(partcmax.df)

```

```{r}
##d. What is the average profit from the sale of Personal Care vs. Cosmetics in each region?
partd.df <- sdf %>% select(X,Region, Total.Profit) %>% 
  group_by(Region,X) %>%
  filter(X == "Personal Care" | X == "Cosmetics") %>%
  summarise(average.profit = mean(Total.Profit), .groups = "drop")

ggplot(partd.df, aes(x = Region, y = average.profit, fill = X)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Profit by Region and Category", x = "Region", y = "Average Profit") +
  scale_fill_manual(values = c("Personal Care" = "#5eb56b", "Cosmetics" = "#855eb5")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
```


```{r}
#e. What is the average profit for each item by region? [5]

parte.df <- sdf %>%
  select(X, Region, Total.Profit ) %>% 
  group_by(Region,X) %>%
  summarise(average.profit2 = mean(Total.Profit), .groups = 'drop')
View(parte.df)

#Turn Each Region into a graph
  
```

##f. In which of the seasons (Spring, Summer, Autumn, Winter) do persons spend the most on:
##i. Cosmetics
##ii. Personal Care [6]

```{r}
getSeason <- function(input.date){
  joined.date<- 100*month(input.date) + day(input.date)
  cuts <- base::cut(joined.date, breaks = c(0,319,0921,1220,1231))
  levels(cuts) <- c("Winter","Spring","Summer","Fall","Winter")
  return(cuts)
}
```


```{r}
partf.df <- sdf %>% 
  select(X, Total.Revenue, Order.Date) %>% 
  filter(X == "Personal Care" | X == "Cosmetics")
partf.df$Season <- getSeason(partf.df$Order.Date)

cosmeticsf.df <-partf.df %>% 
  filter(X == 'Cosmetics') %>%
  group_by(Season,X) %>%
  summarise(TotalRev = sum(Total.Revenue), .groups = "drop") #%>% 
  #slice(which.max(TotalRev))

pie_chart <- ggplot(cosmeticsf.df, aes(x="", y=TotalRev, fill=Season)) + 
  geom_bar(stat="identity", width=1) + 
  coord_polar("y", start=0) +
  scale_fill_manual(values = c("Summer" = "#5eb56b", "Winter" = "#855eb5", "Spring" = "red", "Fall" = "orange")) +
  labs(title = "Total Revenue by for Cosmetic Sales By Season", fill = "Category", x = "", y = "") +
  theme_void() +
  theme(legend.position = "bottom")
print(pie_chart)

PersonalCare.df <- partf.df %>% 
  filter(X == 'Personal Care') %>%
  group_by(Season,X) %>%
  summarise(TotalRev = sum(Total.Revenue), .groups = "drop")

pie_chart2 <- ggplot(PersonalCare.df, aes(x="", y=TotalRev, fill=Season)) + 
  geom_bar(stat="identity", width=1) + 
  coord_polar("y", start=0) +
  scale_fill_manual(values = c("Summer" = "#5eb56b", "Winter" = "#855eb5", "Spring" = "red", "Fall" = "orange")) +
  labs(title = "Total Revenue for Personal Care Sales By Season", fill = "Category", x = "", y = "") +
  theme_void() +
  theme(legend.position = "bottom")
print(pie_chart2)

  
```




##g. Is there a correlation between the units sold for Cosmetics & Personal Care items? Explain
the result. 
```{r}
#corrplot(cor(aq), method = "number")

corr.df <- sdf %>% select(X, Units.Sold) %>% 
  filter(X == "Personal Care") %>% sample_n(6000)
corr2.df <- sdf %>% select(X, Units.Sold) %>%
  filter( X == "Cosmetics") %>% sample_n(6000)
corCosPC <- cor(corr.df$Units.Sold, corr2.df$Units.Sold )
corCosPC

```

##h. Find the monthly average units sold for Personal Care, Cosmetics and Beverages and plot a
##   series graph for 2016. Discuss the resulting visualization. [6]
```{r}
month_dict <- list("1" = "January", "2" = "February", "3" = "March", "4" = "April", "5" = "May", "6" = "June", "7" = "July", "8" = "August", "9" = "September", "10" = "October", "11" = "November", "12" = "December")

parth.df <- sdf %>% 
  filter(X == "Personal Care" | X == "Cosmetics" | X == "Beverages") %>%
  mutate(OrderYear = year(Order.Date))%>%
  #mutate(OrderMonth = month_dict[month(Order.Date)]) %>%
  mutate(OrderMonth = month(Order.Date)) %>%
  
  group_by(X, OrderMonth, OrderYear) %>%
  summarise(avg.sold = mean(Units.Sold), .groups = "drop")
 

timeseriesDf <- parth.df %>% 
  select(everything()) %>%
  filter (OrderYear == 2016)

View(timeseriesDf)
View(parth.df)

timeseriesDf %>% plot(avg.sold~OrderMonth)

ggplot(timeseriesDf, aes(x = OrderMonth, y = avg.sold)) +
  geom_line() +
  labs(title = "Units Sold by Month in 2016",
       x = "Month",
       y = "Units Sold")

```





















